---
title: "Week 4 Part 1- CI's & T-tests"
author: "Jacey Van Wert"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 1. Confidence intervals 

```{r}
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43)

t.test(otter_length)
```

Looking at 95% CI from 34.7-49.8 inch. Means that in that percent of samples that we took from the pop, we would expect to find the population mean ithin the range for the CI presented. 

If we took 100 samples, and found the CI for those samples, then from 95 samples we would expect to find the population mean. 

The mean length of sea otters (N=9) is 42.22 inches. 95 CI is 34.7-49.8 inch.

Default setting is that null hypothesis is mean value equal to zero. Testing null hypothesis that mean otter length = 0 inches. 

## 2. T-test (1-sample t-test)

Can specify the null hypothesis; "?t.test" to get documentation. Define Mu (mean) to change from default setting of mean=zero. 

See a claim: mean otter lenght is 50 inches.

```{r}
otter_test <- t.test(x = otter_length, mu = 50)
otter_test
```

pvalue 0.044, if the population mean really is 50 inches, there is a 4.4% chance that I could've randomly taken a sample from that pop that has a mean that is at least as diff from 50 as my sample mean is in either direction. 

I do have sufficient evidence to reject the null hypothesis. 

## Two-sample t-test

Do not need the same sample size. 

```{r}
desert_bighorns <- c(32, 44, 18, 26, 50, 33, 42, 40)
sierra_bighorns <- c(28, 31, 40, 42, 26, 29, 31)

t.test(x = desert_bighorns, y = sierra_bighorns)

```
Expected confidence interval gives negative because of x and y order.

Difference in means is equal to 0. 

df how far away from 0 (standardized t statistic from how far away means are; different from null hypothesis)

p value tells me that if these samples are drawn from populations with the same mean horn length (null hypoth) then there is almost 48% chance that i couldve taken two ranodom samples from both pops and found samples at least as different as the samples i have here. 

Not enough evidence to reject null hypothesis. 

## End Part 1

## Part 2: Two sample-test

```{r}
compact_suv <- mpg %>% 
  filter(class %in% c("compact", "suv"))
unique(compact_suv$class)
```
### Exploratory data visulization

Exploratory histograms
```{r}
ggplot(data = compact_suv, aes(x=cty))+
  geom_histogram(bins = 15)+
  facet_wrap(~class)
```

Q-Q plot: plots actual values for obs versus theoretical values if perfectly drawn from normal distribution. the more linear, the more normally distributed your sample observations are. 


### Quantile-quantile plot
```{r}
ggplot(data = compact_suv, aes(sample = cty))+
  geom_qq()+
  facet_wrap(~class)

```

Suv looks linear. compact has a few higher than expected gas milage values at high upper tail (some positive skew). 

### Descriptive statistics

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarise( mean_city = mean(cty),
          sd_city = sd(cty),
          n= n())
```

### 2-sample t-test

Create vectors containing observations for each sample (city gas mileage)

Pull because it turns values into vector values. Select keeps values as dataframe. 
```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)


suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)


t.test(x=compact_sample, y=suv_sample)

```
Mean of compact car cty mpg higher than suv. 
95% confidence interval for two sample t test kind of weird.
testing null hypothesis that difference in means is equal to zero- drawn from pops with same exact mean.

the pvalue gives probability (close to zer0) percent chance that i couldve randomly drawn samples with means that are at least this different if they were actually drawn from population with same mean. 

these samples were probably drawn from samples with different means. 